Modeling the Circumnuclear Mediums and Accetion Rates of Quiescent Supermassive Black Holes

### 2014-12-16 ms.tex
mispelled alskeys middlename
updated 2.1

This paper is organized as followed.  In $\S\ref{sec:model}$ we
describe our model, including the sample of galaxies used in our
analysis ($\S\ref{sec:sample}$) and our numerical procedure for
calculating the steady-state hydrodynamic profile of the CNM
($\S\ref{sec:hydro}$).  In $\S\ref{sec:results}$ we describe our
results.  In $\S\ref{sec:applications}$ we describe applicaiton of our
results, including a comparison of our gas profiles to those inferred
from {\it Chandra} observations and to the measured $L_X-\Mbh$
relationship.  In $\S\ref{sec:discussion}$ we discuss our results.  In
$\S\ref{sec:summary}$ we summarize our results.
\section{Model}
\label{sec:model}
\subsection{Galaxy models}
\citealt{LauerFaber+:2007a} used Hubble WFPC2 imaging to measure the
surface brightness profiles for hundred of nearby early type
galaxies. The measured surface brightness profile can generally be
well fit using a Nuker law

\begin{equation}
  I(\xi)=I_b 2^{(\beta-\Gamma)/\alpha} \xi^{-\Gamma} (1+\xi^\alpha)^{-(\beta-\Gamma)/\alpha}, \,\,\,\xi\equiv\frac{r}{r_b}.
\end{equation}
The Nuker profile is a broken power law which transitions from an
inner power law slope, $\Gamma$, to an outer power law slope, $\beta$,
at a break radius, $\rb$.

Assuming spherical symmetry this corresponds to a stellar density
$\rhostar \sim r^{-1-\Gamma}$ for $r \ll \rb$ and $\rhostar\sim
r^{-1-\beta}$ for $r \gg \rb$. Thus, we adopt a broken power law for 
$\rhostar$

\begin{align}
\rho_\star = 
\begin{cases}
\rho_\star(\rsoi) \left(r/\rsoi\right)^{-1-\Gamma} & r \leq r_b\\
\rho_\star(r_b) \left(r/r_b\right)^{-1-\beta} & r > r_b,
\end{cases}
\end{align}

where $\rho_\star(\rsoi)$ is the stellar density at the black hole
sphere of influence, $\rsoi$. $\rho_\star(\rsoi)$ is chosen so that
the stellar mass inside $\rsoi$, $\Mstar(\rsoi)$, is equal to the
black hole mass.  $\rsoi=G \Mbh/\sigma^2$. Using the $\Mbh-\sigma$ relation
from \citealt{Gultekin+09}

 \begin{align}
M_{\bullet} \simeq 2\times 10^{8}\left(\frac{\sigma}{200\,{\rm
      km\,s^{-1}}}\right)^{5.1}M_{\odot}, 
\label{eq:Msigma}
\end{align}

we obtain $\rsoi \approx 14 M_{\bullet,8}^{0.6}$. In order to
completely specify the galaxy model we must choose $\Mbh$, $\Gamma$, $r_b$,
and $\beta$. We compute models for three different models of
$\Mbh$: $10^6$, $10^7$, $10^8 \Msun$. In the
\citealt{LauerFaber+:2007a} sample the distribution of the $\Gamma$s
is bimodal, with a concentration of core galaxies with $\Gamma \lsim
0.3$ and a concentration of cusp galaxies with $\Gamma \gsim
0.7$. We calculate models for $\Gamma=0.8$ and $\Gamma=0.1$.

We are interested in the gas flow on scales smaller than
the break radius $\rb$, which typically occurs at a radius of 100s of
parsecs. Thus, we fix $\beta$ to be 2 {\bf AG:I have not yet checked this
  directly via numerical experiments}.  The presence of $\rb$ is 
necessary for to obtain a converged steady state for some regions of
our parameter space. For cusp galaxies, $r_b$ is a few hundred parsecs
independently of $\Mbh$. We adopt $\rb=100$ pc for our cusp
galaxies ($\Gamma=0.8$), while for our core galaxy models we use the
scaling relation...{\bf AG: Need to figure out normalization for
  scaling relation from Lauer. Fixed rb at a given Mbh only has a
  factor of a few effect on rs--this will be discussed in the section
  on numerical solutions.}

% \subsection{Galaxy Sample}

% \label{sec:sample}
% %AG:It is unclear whether if/how the Nuker profile will show up in
% %this parametrized study.
% {\bf BDM: we may show plot of NUKER masses, $\Gamma$'s, etc here to (1) motivate the parametric study; (2) in individual cases we may want to compare solutions to specific galaxies.}

% Our sample includes 61 early-type galaxies from
% \citetalias{WangMerritt:2004a}, as represent a subset of a larger
% sample of \citet{FaberTremaine+:1997a} with Hubble WFPC2 imaging.  We
% focus on elliptical galaxies because we are interested in isolating
% SMBHs fed by the steady mass loss from the relatively old stellar
% population, instead of fresh gas flowing into the sphere of influence
% from larger scales.

% The surface brightness of each galaxy in the sample is parameterized by a Nuker Law (WM04; their Table 1),
% \begin{equation}
%   I(\xi)=I_b 2^{(\beta-\Gamma)/\alpha} \xi^{-\Gamma} (1+\xi^\alpha)^{-(\beta-\Gamma)/\alpha}, \,\,\,\xi\equiv\frac{r}{r_b}.
% \end{equation}
% The Nuker profile is a broken power law which transitions from an inner power law slope, $\Gamma$, to an outer power law slope, $\beta$, at a break radius, $\rb$.

% Given $I(\xi)$ we follow the procedure outlined by
% \citetalias{WangMerritt:2004a} to compute the spherically-symmetric
% stellar density profile $\rhostar(r)$ and stellar mass profile
% $M_{\star}(r) = 4\pi \int \rhostar r^{2}dr$, both of which are used in
% our model.  For reference, note that stellar density $\rhostar\sim
% r^{-1-\Gamma}$ for $r \ll \rb$ and $\rhostar\sim r^{-1-\beta}$ for $r
% \gg \rb$.  Since $\Gamma\approx 1$ for ``cusp" galaxies, while
% $\Gamma\approx 0$ for ``core" galaxies, at small radii $\ll \rb$ we
% thus have $\rhostar \propto r^{-\alpha}$ and $M_{\star} \propto
% r^{\beta}$ with $\beta = 3-\alpha \approx 1-2$.  For each galaxy the
% SMBH mass $\Mbh$ is determined from the $\Mbh-\sigma$ relationship.
% %%AG-ugh I just took the Mbh from Wang & Merritt, which I guess is
% %%calculated from an outdated M-sigma. It would be bizarre to use 2
% %%different relations.

% AG:Is there a systematic dependence of beta on galaxy type$.

%% AG--some discussion of alternate parameterizations.  My intuition
%% is that for our purposes the parameterization should not
%% matter. But some discussion of alternate possibilities could
%% prevent Alistar Graham from yelling at us.

@@

The source term in the continuity equation (equation~\ref{eq:drhodt}),
\begin{align}
  q=\frac{\eta \rhostar}{\th},
\label{eq:q}
\end{align}
represents mass input from stellar winds, which we parameterize in
terms of the fraction $\eta$ of the stellar density $\rhostar$ being
recycled into gas on the Hubble time $\th = 1.4 \times 10^{10}$
years. $\eta\simeq 0.02 \left(\frac{t}{th}\right)^{-1.3}$ 
(see, e.g., \citealt{Ciotti+91})\footnote{Note that 
\citealt{Ciotti+91} gives the overall mass return rate from evolved
stars as a function of B-band luminosity. We give the mass return rate
per unit volume. However, our expressions are equivalent.}

@@

\section{Analytic model for dependence of wind heating $\vwO^{\star}$ on stellar age}
\label{app:windheat}
\input{windheat}

>>>>

This paper is organized as followed.  In $\S\ref{sec:model}$ we
describe our model, including the sample of galaxies used in our
analysis ($\S\ref{sec:sample}$) and our numerical procedure for
calculating the steady-state hydrodynamic profile of the CNM
($\S\ref{sec:hydro}$).  In $\S\ref{sec:results}$ we describe our
results.  In $\S\ref{sec:applications}$ we describe applicaiton of our
results, including a comparison of our gas profiles to those inferred
from {\it Chandra} observations and to the measured $L_X-\Mbh$
relationship.  In $\S\ref{sec:discussion}$ we discuss our results.  In
$\S\ref{sec:summary}$ we summarize our results.
\section{Model}
\label{sec:model}
\subsection{Galaxy models}
\citet{LauerFaber+:2007a} use Hubble WFPC2 imaging to measure the
radial surface brightness profiles for hundreds of nearby early type galaxies. The measured surface brightness profile is generally well-fit by a ``Nuker" law parameterization:
\begin{equation}
  I(\xi)=I_b 2^{(\beta-\Gamma)/\alpha} \xi^{-\Gamma} (1+\xi^\alpha)^{-(\beta-\Gamma)/\alpha}, \,\,\,\xi\equiv\frac{r}{r_b}.
\end{equation}

The Nuker profile is a broken power law which transitions from an inner power law slope, $\Gamma$, to an outer power law slope, $\beta$, at a break radius, $\rb$.  Assuming spherical symmetry, this corresponds to a stellar density $\rhostar \sim r^{-1-\Gamma}$ for $r \ll \rb$ and $\rhostar\sim r^{-1-\beta}$ for $r \gg \rb$, thus motivating us to adopt a broken power-law profile
\begin{align}
\rho_\star = 
\begin{cases}
\rho_\star(\rsoi) \left(r/\rsoi\right)^{-1-\Gamma} & r \leq r_b\\
\rho_\star(r_b) \left(r/r_b\right)^{-1-\beta} & r > r_b,
\end{cases}
\end{align}
where $\rho_\star(\rsoi)$ is the stellar density at the radius of the black hole sphere of influence, 
\be
\rsoi \simeq G \Mbh/\sigma^2 \approx 14 M_{\bullet,8}^{0.6},
\label{eq:rsoi}
\ee
and the final equality in (\ref{eq:rsoi}) employs a recent version of the $\Mbh-\sigma$ relationship (\citealt{Gultekin+09}),
 \begin{align}
M_{\bullet} \simeq 2\times 10^{8}\left(\frac{\sigma}{200\,{\rm
      km\,s^{-1}}}\right)^{5.1}M_{\odot}.
\label{eq:Msigma}
\end{align}

A galaxy model is fully specified by four parameters: $\Mbh$, $\Gamma$, $r_b$,
and $\beta$.  We compute models for three different black hole masses, $\Mbh = 10^6$, $10^7$, $10^8 \Msun$.  The distribution of $\Gamma$ in the \citet{LauerFaber+:2007a} sample is bimodal, with a concentration of ``core" galaxies with $\Gamma \lsim 0.3$ and a concentration of ``cusp" galaxies with $\Gamma \gsim 0.7$.  We bracket these possibilities by considering models with $\Gamma=0.8$ and $\Gamma=0.1$.  Since we are generally interested in properties of the gas flow on radial scales $\ll r_b$, we fix $\beta = 2$ (a typical value), an assumption to which most of our results is not sensitive.  However, the presence of $\rb$ is necessary for to obtain a converged steady state for some regions of our parameter space. We adopt $\rb=100$ pc for cusp galaxies ($\Gamma=0.8$), while for core galaxy models we employ the scaling relation $\rb = $, motivated again by the \citet{LauerFaber+:2007a} sample. ({\bf BDM: fill in!}).  Though simplistic, our results are not sensitive to our assumption about the value of $\rb$.

@@

The source term in the continuity equation (eq.~[\ref{eq:drhodt}]),
\begin{align}
  q=\frac{\eta \rhostar}{\th},
\label{eq:q}
\end{align}
represents mass input from stellar winds, which we parameterize in
terms of the fraction $\eta$ of the stellar density $\rhostar$ being
recycled into gas on the Hubble time $\th = 1.4 \times 10^{10}$
years. $\eta\simeq 0.02 \left(\frac{t}{th}\right)^{-1.3}$ 
(see, e.g., \citealt{Ciotti+91})\footnote{Note that 
\citealt{Ciotti+91} gives the overall mass return rate from evolved
stars as a function of B-band luminosity. We give the mass return rate
per unit volume. However, our expressions are equivalent.}

@@

%\section{Analytic model for dependence of wind heating $\vwO^{\star}$ on stellar age}
%\label{app:windheat}
%\input{windheat}
